Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-08-10 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
dealt with comments in the supplementaries

\subsection{Effect of the starting tree on Bayesian inference}
We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference to decrease the computing time needed.
To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random starting tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
We tested this effect on five trees with the five levels of missing data (i.e. first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100 (i.e. 0\% ``missing''); second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
We also compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
% NC: Why the regressions? What are they telling you? Why simple and nested?
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).

\begin{table}[ht]
\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.} % NC: I'm not sure what these two tables are telling you. Clarify. The caption says nothing about the random starting trees.
\label{Tab_lm_res}
\centering
\begin{tabular}{rlcccc}
  \hline
 & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
RF & (Intercept) & 0.8100 & 0.0122 & 66.16 & \textbf{0.0000} \\ 
   & starting:true & 0.0019 & 0.0173 & 0.11 & 0.9125 \\
Tr & (Intercept) & 0.4424 & 0.0418 & 10.59 & \textbf{0.0000} \\ 
   & starting:true & 0.0159 & 0.0591 & 0.27 & 0.7887 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics for each missing data parameter using nested linear regression.}
\label{Tab_lmList_res}
\centering
\begin{tabular}{rccccc}
  \hline
 & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
RF & 0\%  & 0.02 & 0.02 & 1.02 & 0.31 \\ 
   & 10\% & -0.01 & 0.02 & -0.50 & 0.62 \\ 
   & 25\% & 0.00 & 0.02 & 0.00 & 1.00 \\ 
   & 50\% & 0.00 & 0.02 & 0.03 & 0.98 \\ 
   & 75\% & -0.00 & 0.02 & -0.03 & 0.98 \\
Tr & 0\%  & 0.09 & 0.08 & 1.06 & 0.29 \\ 
   & 10\% & -0.00 & 0.08 & -0.05 & 0.96 \\ 
   & 25\% & 0.00 & 0.08 & 0.01 & 0.99 \\ 
   & 50\% & -0.00 & 0.08 & -0.04 & 0.97 \\ 
   & 75\% & -0.00 & 0.08 & -0.00 & 1.00 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

\subsection{Effect of the starting tree on Bayesian inference}
We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference to decrease the computing time needed.
To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random starting tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
We tested this effect on five trees with the five levels of missing data (i.e. first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100 (i.e. 0\% ``missing''); second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
We used a two-way ANOVA to test any significant effect of the starting tree (``true'' or random) on the normalised Robinson-Foulds and Triplets metrics.
% NC: Why the regressions? What are they telling you? Why simple and nested?
% TG: Ok, made it totally plain easier. Anova.
The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (table ~\ref{Tab_lm_res}).

\begin{table}[ht]
\caption{Test of the effect of using either a random tree or the ``true'' tree as a starting tree on two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using a two-way ANOVA.}
\label{Tab_lm_res}
\centering
\begin{tabular}{rlcccccc}
  \hline
 metric & terms & Df & Sum Sq & Mean Sq & F value & Pr($>$F) \\ 
  \hline
RF & starting    & 1 & 0.00 & 0.00 & 0.01 & 0.9125 \\ 
   & Residuals   & 198 & 2.97 & 0.01 &  &  \\ 
Tr & starting    & 1 & 0.01 & 0.01 & 0.07 & 0.7887 \\ 
   & Residuals   & 198 & 34.57 & 0.17 &  &  \\ 
   \hline
\end{tabular}
\end{table}

### 2015-08-10 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
dealt with comments in the supplementaries

\subsection{Differences between the ``true'' and the ``best'' trees.}
% NC: Need a line or two to explain the motivation of this analysis. Why did you do it? (not just because reviewers asked you to!)
We compared our ``true'' trees (i.e. the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e. the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred
 the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer % NC: again
phylogenies from this particular matrices (i.e. small sized and generated using specific algorithms).
This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.

>>>>

\subsection{Differences between the ``true'' and the ``best'' trees.}
% NC: Need a line or two to explain the motivation of this analysis. Why did you do it? (not just because reviewers asked you to!)
% TG:  How's that?
In our simulation protocol, we used the ``true'' tree to generate the molecular characters and the morphological characters for the living and fossil taxa (i.e. the ``complete'' matrix).
The following analysis measures the performance of our parameter and algorithms choices to generate the ``complete'' matrix.
%
To asses the performance of our simulation protocol, we compared our ``true'' trees (i.e. the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e. the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred TG: no, it's just the matrices that are not inferred anymore.
 the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer phylogenies from this particular matrices (i.e. small sized and generated using specific algorithms).
This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.
